### 2023-03-15 ms1_data_analysis.Rmd
Simple model for SLA added to the summary.

---
title: "Data Analysis"
author: "Piotr Szefer"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, results='hide',warning=FALSE, message=FALSE)

```

## Experimental design

```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE}
# The source file loads the data preparation script, where the raw data is read and summarised for further analyses
source('code/log_ratio_extraction.R')
```

We collected data on biomass (total of `r round(sum(BIOMASS$tot_bio), 0)` kg), density, and plant traits: area lost to herbivores, water content, leaf dry matter content and specific leaf area for `r length(unique(data$unified_names))` woody plant species from three tropical forest sites on elevations spanning `r 1900 - 200` m from 200 to 1900 m a.s.l. Low elevation is Wanang (200 m a.s.l.), Mid-elevation is Numba (700 m a.s.l.), and High elevation is Yawan (1900 m a.s.l.). We refer to the sites as Low, Mid- and High elevation throughout the text.

Figure 1. Median, 1st, second quadrilles, 1.5 IRQ distance and outliers in biomass, richness diversity and density at each location/elevation, but some Yawan sites were extremely productive.

```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap= "Figure S1 Characteistics"}

library(ggplot2)
library(dplyr)

elev.bind <- rbind(BIOMASS %>% dplyr::select(site, garden, treatment, tot_bio), 
                   RICHNESS %>% dplyr::select(site, garden, treatment, sp_no), 
                   RICHNESS %>% dplyr::select(site, garden, treatment, simpson), 
                   DENSITY %>% dplyr::select(site, garden, treatment, tot.stem.no))

findoutlier <- function(x) {
  return(x < quantile(x, .25) - 1.5*IQR(x) | x > quantile(x, .75) + 1.5*IQR(x))
}

elevdata <- elev.bind %>% 
  mutate(value = ifelse(!is.na(tot_bio), tot_bio, 
                        ifelse(!is.na(sp_no), sp_no, 
                               ifelse(!is.na(simpson), simpson, tot.stem.no)))) %>%
  mutate(descriptor = ifelse(!is.na(tot_bio), "Total biomass [kg]", 
                        ifelse(!is.na(sp_no), "Number of species", 
                               ifelse(!is.na(simpson), "Simpson's diversity", "Abundance")))) %>%
  dplyr::select(site, garden, treatment, value, descriptor) %>%
  mutate(fsite = factor(site, levels = c("wanang", "numba","yawan"))) %>%
  arrange(descriptor, fsite, treatment, garden)

elevdataT <- elevdata %>% 
  mutate(valueTrans = ifelse(descriptor == "Simpson's diversity", 
                             value, log(value)))

outliers <- c()
for (desc in unique(elevdataT$descriptor)) {
  for (st in unique(elevdataT$site)) {
    for (trt in unique(elevdataT$treatment)) {

      print(paste(desc,st,trt))

      outliers <- c(outliers,findoutlier(pull(elevdataT %>% filter(descriptor == desc,
                          site == st,
                          treatment == trt), valueTrans)))

    }

  }
}
elevdataT$outlierBool <- outliers 
elevdataT <- elevdataT %>% mutate(outlier = ifelse(outlierBool, garden, ""))

ggplot(elevdataT, aes(y = valueTrans, x = fsite, label = garden))+
  geom_boxplot()+
  geom_text(aes(label=outlier), hjust = 1.2, vjust = -0.0005, na.rm=TRUE)+
  facet_grid(vars(descriptor), vars(treatment), scales = "free")+
  theme_bw()

```

Looking at the control plots Numba (the mid-elevation in our study) has the highest abundance and number of species. Predictably productivity decreases with elevation. Control plots in Numba are more variable than in other two elevations. Yawan has one site with lowest abundance (garden number seven).However no significant differences in variance were detected (Levene's and Bartellet's tests)

## General effects of biotic factors

We calculated log-response ratios for plant biomass, species richness, species diversity and stem density to be able to compare the effect sizes at individual elevations.

Table 1. Test results for over-dispersion all descriptors at three studied elevations.

```{r, echo=FALSE, results='show', fig.show='all', warning=FALSE, message=FALSE}
sjPlot::tab_df(lrr_test_df)
```

```{r,  echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE}

# Results of a GLS model with over-dispersion vs LM model for density LRR in the predator exclusion treatment.
trtdata <-  trtdata  <- DENSITY.LR %>%
    filter(treatment == 'p')

trtlm <- lm(lratio ~ 0 + site, data = trtdata)

trtgls <- nlme::gls(lratio ~ 0 + site, data = trtdata,
                 weights = varIdent(form = ~1|site))

summary(trtlm)
summary(trtgls)

aovtrt <- anova(trtlm)

testStat <- aovtrt$`F-value`
testP <- aovtrt$`p-value`

# Tukey post-hoc
emm <- emmeans(object = trtlm,
               specs = "site",
               adjust = "mvt")

mod_means <- multcomp::cld(object = emm,
                           Letters = letters)$.group

zero.dif.lm <- ifelse(summary(trtlm)$coefficients[,4] > 0.05, "","*")
zero.dif.gls <- ifelse(summary(trtgls)$tTable[,4] > 0.05, "","*")

lmres <- as_tibble(summary(trtlm)$coefficients)
lmres <- rename(lmres, Estimate = 'Estimate', 
               `Standard Error` = `Std. Error`, 
               `T-value` =`t value`,
               `P-value` = `Pr(>|t|)`)

glsres <- as_tibble(summary(trtgls)$tTable)
glsres <- rename(glsres,
                 Estimate = 'Value', 
                 `Standard Error` = "Std.Error", 
                 `T-value` = "t-value", 
                 `P-value` = 'p-value')

model_comp <- rbind(lmres, glsres)
model_comp$Site <- rep(c("Wanang", "Numba", "Yawan"), 2)
model_comp$Model <- rep(c("LM", "GLS"), each = 3)

```

Table S2. Two models testing difference of LRR of predator exclosures from zero at each elevation.

```{r}
sjPlot::tab_df(model_comp)
```

There is generally little evidence for the over-dispersion in the LRR values at individual elevations. Only woody stem density for the predator exclusion showed some evidence of over-dispersion. Including this in statistical modle (using 'varIdent(form = ~1|site)' as weights in the GLS function) showed significant difference of LRR in Yawan, that was not observed for LM model.

```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap= "Figure 1. Magnitude and direction of fungi, insects and their predators on biomass, richness, diversity and density of woody plants along elevation gradient: Wanang (200 m a.s.l.); Numba (750 m a.s.l.); Yawan (1900 m a.s.l.). Log response ratio is a log-ratio of a given descriptor value from plots where a biotic factor was present to where it was absent. Grey points represent empirical values. Letters indicate statistically significant ($\\alpha$ = 0.05) differences for pairwise comparisons between elevations within a treatment-descriptor combination with Tukey correction for multiplicity. For better readability letters are given only in cases where any of the mean were either different form each other or different from zero. Zero effect are indicated with a dashed line. Stars in the upper index indicate significant ($\\alpha$ = 0.05) differences of the effect mean from zero (significance of the effect at a given site)", fig.width= 8, fig.height=8}

# bio, dens, div, rich
panel_data <- data.frame()

dats <- list(biomass = BIOMASS.LR, 
             denssity = DENSITY.LR, 
             diversity = DIVERSITY.LR, 
             richness = RICHNESS.LR)

# Stack the data and rename descriptor names
for (file in names(dats)){
  print(file)
  dat <- dats[[file]]
  dat$descriptor <- ifelse(grepl("bio", file), "Biomass",
                           ifelse(grepl("dens", file), "Woody plant density",
                                  ifelse(grepl("div", file),"Diversity", "Richness")))
  dat <- dat[, c("site", "garden", "treatment", "lratio","descriptor")]
  panel_data <- rbind(dat, panel_data)
}

# Make a renamed factor variable for the treatments

panel_data$ftreatment <- panel_data$treatment

# Fix treatment names
panel_data$ftreatment[panel_data$ftreatment == "i"] <- "Insects"
panel_data$ftreatment[panel_data$ftreatment == "p"] <- "Birds, bats and ants"
panel_data$ftreatment[panel_data$ftreatment == "f"] <- "Fungi"
panel_data$ftreatment[panel_data$ftreatment == "h"] <- "Increased herbivory"
panel_data$ftreatment <- factor(panel_data$ftreatment,
                           levels = c("Fungi",
                                      "Insects",
                                      "Birds, bats and ants",
                                      "Increased herbivory"
                                      ))

# Set the transparency value
aph <- 0.5

# Create a factor variable for sites
panel_data$fsite <- factor(panel_data$site, labels = c("Wanang",
                                      "Numba",
                                      "Yawan"))

# Display panel plot
ggplot(panel_data, aes(x = fsite, y = lratio))+
  geom_jitter(width = 0.1, alpha = 0.1, size = 2)+

  # Group means
  stat_summary(fun = mean, na.rm = T,
               geom = "point", size = 2,
               alpha = aph)+

  stat_summary(fun.data = "mean_cl_boot",
               alpha = aph) +

  # Annotations
  geom_text(data = panel_data %>%
              group_by(ftreatment, fsite) %>%
              summarise(mlr = mean(lratio)),
            aes(x = fsite, y = mlr + 1.5, group = ftreatment,
                label = lrr_panel_labs),
            size = 3)+

  theme_bw()+
  theme(strip.text = element_text(size =5.5),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1))+
  geom_hline(yintercept = 0, lty = 2)+
  ylab("Log response ratio")+xlab("")+
  facet_grid(descriptor ~ ftreatment)

```

## Impact of treatments

Positive effect of pathogenic fungi on richness was found at low and high elevations (Fig. 1) and on density at low elevation. Insects reduced biomass only at the mid-elevation, increased diversity at high and richness at high and low elevation. At high elevation predator removal caused an increase in diversity, richness and density of woody plants.Increase in generalist herbivore abundance reduced biomass and woody plant density at low elevation and richness at low and high elevations.

```{r}

# Check whether the RICHNESS.LR for predators is correct.
# trtdata <-  trtdata  <- RICHNESS.LR %>%
#     filter(treatment == 'p')

# trtlm <- lm(lratio ~ 0 + site, data = trtdata)

# summary(trtlm)

# Tukey post-hoc
# emm <- emmeans(object = trtlm,
#                specs = "site",
#                adjust = "mvt")

# mod_means <- multcomp::cld(object = emm,
#                            Letters = letters)$.group

```

## LRR correlations

Differences of LRR averages from zero may suggest that the effects are random, generally weak and noisy. We now look at the LRR correlation for different treatments to probe some possible mechanisms of the biotic factors' action.

```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap= "Figure 2. ...", fig.width= 10, fig.height= 10}

# library(dplyr)
# library(ggplot2)
# library(GGally)
# library(lme4)
# library(lmerTest)
# library(ggpmisc)
# library(tibble)

# LRR correlation plot

ftreats <- unique(panel_data$ftreatment)

fdat <- panel_data %>%
  filter(ftreatment == ftreats[1])%>%
  arrange(descriptor, site, garden)

hdat <- panel_data %>%
  filter(ftreatment == ftreats[2])%>%
  arrange(descriptor, site, garden)

idat <- panel_data %>%
  filter(ftreatment == ftreats[3])%>%
  arrange(descriptor, site, garden)

pdat <- panel_data %>%
  filter(ftreatment == ftreats[4])%>%
  arrange(descriptor, site, garden)

# chdat <- panel_data %>%
#   filter(ftreatment == ftreats[5])%>%
#   arrange(descriptor, site, garden)

# Create one dataset
pdpairs <- panel_data %>%
  group_by(descriptor, site, garden) %>%
  summarise() # Summarise based on what??????

# Add columns with LRRs
pdpairs$f <- fdat$lratio
pdpairs$h <- hdat$lratio
pdpairs$i <- idat$lratio
pdpairs$p <- pdat$lratio

biolrr <- pdpairs %>% filter(descriptor == "Biomass")
biocont <- BIOMASS  %>% filter(treatment == "c") %>%
  arrange(as.character(site), garden)

cpdata <- data.frame(descriptor = rep(pdpairs$descriptor,2),
                     site = rep(pdpairs$site,2),
                     garden = rep(pdpairs$garden),
                     x = c(pdpairs$h,
                           pdpairs$p),
                     y = rep(pdpairs$i, 2),
                     response = rep(c("Elevated herbivory effect",
                                      "Top predators effect"), 
                                    each = length(pdpairs$descriptor)))

lrr.cor.test.data <- c()

for (resp in unique(cpdata$response))   {
  for (desc in unique(cpdata$descriptor)){
    for (loc in unique(cpdata$site))     {

      print(paste(resp, desc,loc))

      lrrtd <- cpdata %>%
        filter(response == resp,
               descriptor == desc,
               site == loc)

      # np.test.p <- cor.test(lrrtd$x, lrrtd$y, method = c('pearson'))$p.value
      # Test for noramlity
      print(shapiro.test(lrrtd$x))
      print(shapiro.test(lrrtd$y))

      lm.test.p <- summary(lm(lrrtd$y~lrrtd$x, data = lrrtd))$coefficients[2,4]

      print(lm.test.p)

      lrr.res <- data.frame(response = resp,
                            descriptor = desc,
                            site = loc,
                            non.param.p = ifelse(lm.test.p<0.05,"signif","ns"))
      lrr.cor.test.data <- rbind(lrr.cor.test.data, lrr.res)

    }
  }
}

lrr.cor.test.data -> lctd

cpdata$sign <- ""

sign.dat <- lctd[lctd$non.param.p == "signif",]

for(row in 1:dim(sign.dat)[1]){
  resp <- sign.dat[row, ]$response
  desc <- sign.dat[row, ]$descriptor
  loc <- sign.dat[row, ]$site

  c1 <- cpdata$response == resp
  c2 <- cpdata$descriptor == desc
  c3 <- cpdata$site == loc

  cpdata[c1 & c2 & c3, ]$sign <- "signif"
}

ggplot(cpdata, aes(x = x, y = y)) +
  geom_point(size = 2, alpha = 0.4)+
  stat_smooth(data = cpdata %>%
                filter(sign == "signif"),
              method ="lm")+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
  ylab("Effect size of insects")+
  xlab("Effect size")+
  theme_bw()+
  facet_grid(response*descriptor~site)

```

#### OBSERVATIONS

Correlation between LRR of insect effects were not very common and were observed in `r 9/(3*8) * 100` % of cases. Correlations were more often found in Yawan (five out of eight studied correlations) than in Numba (four out of eight) and no significant correlation was found in Wanang. If we also include correlations between predator exclusion (P) and P+A this ratio gets significantly higher for Yawan (9 out of 12 correlations).

All observed correlations of the top-predators effects were positively correlated with insect effects, whereas all effects of herbivore addition were negatively correlated with the insect effects.

#### EXPLANATION

If a plant community has a potential to be strongly negatively affected by insects (negative LRR of insects, strong top-down control), then the P+A would result in weaker or positive effects of additional generalist herbivores. Top down control seem to be stronger in the highlands and we see more frequent correlation of the LRR's. (Model where Intra-guild predators are being strongly limited by top predators.)

What exactly happened after addition can however only be speculated here. In case of a  strong bottom-up control should not result in any sort of correlation like in Wanang. What I think happened was that spiders could be more active in the exclosures which resulted in their effects being similar to that of a simple exclusion.

A few aspects of insect effects on plants can be important here: 
If community is saturated with herbivores (Konno) then addition may have negative results.
Herbivore resource limitation type (relative or absolute). Feeding mode change, adaptive trade off, risk avoidance. All these can only be speculated.

Generalists are worse in consuming any individual plant biomass, therefore they may have positive effects on plants, when natural insect communities have negative effects. It is possible that addition of herbivores cause som disruption in the functioning of the insect community effect on vegetation.
There were no correlation between factors’ effects on vegetation at the low elevation. Herbivore addition and removal were not correlated as expected by the strong bottom-up control.

Simulation showed that it cannot be simply an effect of the effect-abundance relationship between abundance of insects and their effect (quadratic and linear function result in positive relationship).

Top-down effects of predators and their positive correlation with insect effects can be explained well by the bottom-up control of insects by plants. It was also assumed in the above simulations. Empirically we do not observe any correlation between insect abundance and strength of the effect (lrr_simualtions.R at the end). There is also no correlation between richness of a community and its LRR for any of the treatments (lrr_simulations.R).

#### SOME DISCLAIMERS

It is noteworthy that plots where herbivores were added were also covered with nets to prevent increased predation, wich is known to be correlated with increased leaf damage (Katka). Therefore it cannot be interpreted directly as a simple increase in herbivory pressure even though this is what should be happening assuming simple chain model and top-down control by top-predators. 
We do not know the faith of the added predators. It is most likely that they did not cause a permanent increase in the species abundance (graph here). However, herbivore damage in general was...

### Predator exclusion and predator exclusion with insect addition

The hypothesis that I have stated above, about increased activity of spiders in exclosures is here being investigated further. I want to see whether the addition of herbivores has a positive correlation with the regular exclosures. Positive correlation would suggest that the herbivore addition had little to no effect, and that spiders responding to a lack of predators would control the herbivore population even with boosted abundances.

```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap= "Figure 2. Log-Response Ratio correlations for insects and fungi effects on bimoass, diversity adn density of woody plants"}

fidata <- data.frame(descriptor = pdpairs$descriptor,
                     site = pdpairs$site,
                     garden = rep(pdpairs$garden),
                     x = pdpairs$p,
                     y = pdpairs$h)

lrr.cor.test.data <- c()

# for (resp in unique(fidata$response))   {
  for (desc in unique(fidata$descriptor)){
    for (loc in unique(fidata$site))     {

      print(paste(desc,loc))

      lrrtd <- fidata %>%
        filter(descriptor == desc,
               site == loc)

      # Test for normality
      print(shapiro.test(lrrtd$x))
      print(shapiro.test(lrrtd$y))

      lm.test.p <- summary(lm(lrrtd$y~lrrtd$x, data = lrrtd))$coefficients[2,4]

      print(lm.test.p)

      lrr.res <- data.frame(descriptor = desc,
                            site = loc,
                            non.param.p = ifelse(lm.test.p<0.05,"signif","ns"))
      lrr.cor.test.data <- rbind(lrr.cor.test.data, lrr.res)

    }
  }
# }

lrr.cor.test.data -> lctd

fidata$sign <- ""

sign.dat <- lctd[lctd$non.param.p == "signif",]

for(row in 1:dim(sign.dat)[1]){
  # resp <- sign.dat[row, ]$response
  desc <- sign.dat[row, ]$descriptor
  loc <- sign.dat[row, ]$site

  # c1 <- cpdata$response == resp
  c2 <- fidata$descriptor == desc
  c3 <- fidata$site == loc

  fidata[c2 & c3, ]$sign <- "signif"
}

ggplot(fidata, aes(x = x, y = y)) +
  geom_point(size = 2, alpha = 0.4)+
  stat_smooth(data = fidata %>%
                filter(sign == "signif"),
              method ="lm")+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
  xlab("Predator removal")+
  ylab("Predator removal and insect addition")+
  theme_bw()+
  facet_grid(descriptor~site)

```

From the figure we see that there is usually no correlation in Wanang and Numba but it is always present in Yawan. Moreover it is not positive but negative. This means that when we observer negative effect of predator removal on biomass we can expect to have positive effect of insect addition to the exclosure. In cases where the effect of predators is positive, we would get a negative effect of insect addition.

### Fungi vs insect correlation
No correlation should be observed in case of compensatory effect
In case of species richness only effects of insect, herbivore addition and herbivore removal were correlated. Herbivore addition and removal were negatively correlated as expected. If there is a compensatory effect there should be no correlation (because when we remove one factor the other takes over and equalize the effect). Positive correlation rules out compensation. A positive correlations would suggest bottom-up control.

```{r, echo=FALSE, results='hide', fig.show='all', warning=FALSE, message=FALSE, fig.cap= "Figure 2. Log-Response Ratio correlations for insects and fungi effects on bimoass, diversity adn density of woody plants"}

fidata <- data.frame(descriptor = pdpairs$descriptor,
                     site = pdpairs$site,
                     garden = rep(pdpairs$garden),
                     x = pdpairs$i,
                     y = pdpairs$f)

lrr.cor.test.data <- c()

# for (resp in unique(fidata$response))   {
  for (desc in unique(fidata$descriptor)){
    for (loc in unique(fidata$site))     {

      print(paste(desc,loc))

      lrrtd <- fidata %>%
        filter(descriptor == desc,
               site == loc)

      # Test for normality
      print(shapiro.test(lrrtd$x))
      print(shapiro.test(lrrtd$y))

      lm.test.p <- summary(lm(lrrtd$y~lrrtd$x, data = lrrtd))$coefficients[2,4]

      print(lm.test.p)

      lrr.res <- data.frame(descriptor = desc,
                            site = loc,
                            non.param.p = ifelse(lm.test.p<0.05,"signif","ns"))
      lrr.cor.test.data <- rbind(lrr.cor.test.data, lrr.res)

    }
  }
# }

lrr.cor.test.data -> lctd

fidata$sign <- ""

sign.dat <- lctd[lctd$non.param.p == "signif",]

for(row in 1:dim(sign.dat)[1]){
  # resp <- sign.dat[row, ]$response
  desc <- sign.dat[row, ]$descriptor
  loc <- sign.dat[row, ]$site

  # c1 <- cpdata$response == resp
  c2 <- fidata$descriptor == desc
  c3 <- fidata$site == loc

  fidata[c2 & c3, ]$sign <- "signif"
}

ggplot(fidata, aes(x = x, y = y)) +
  geom_point(size = 2, alpha = 0.4)+
  stat_smooth(data = fidata %>%
                filter(sign == "signif"),
              method ="lm")+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
  ylab("Effect size of insects")+
  xlab("Effect size of fungi")+
  theme_bw()+
  facet_grid(descriptor~site)

```

### COMMUNITY COMPOSITION

At each elevation insect exclusion significantly affected the community composition [but only when all treatments were included, in pairwise comparisons, i.e., C vs I the composition was not changed significantly]. Fungicide treatment significantly affected the full plant community composition only at the highest elevation (Fig, 3; Appendix I, Tab. S1).

```{r}

library(vegan)
library(cowplot)

# 1. Load the data
COMMDATA <- t(COMMDATA)
COMMVARS

# Select only i,c,p,f
COMM.SEL <- COMMDATA[grepl("i|c$|p|f|_h", rownames(COMMDATA)), ]
VARS.SEL <- COMMVARS[grepl("i|c$|p|f|_h", COMMVARS$plotid), ]

treats.to.consider <- c("i","c","h","f","p")

# Dummy-coding of the treatments
COMMVARS <- COMMVARS %>%
  mutate(f = ifelse(treatment == "f", 1, 0),
         h = ifelse(treatment == "h", 1, 0),
         p = ifelse(treatment == "p", 1, 0),
         c = ifelse(treatment == "c", 1, 0),
         i = ifelse(treatment == "i", 1, 0))

# RDA at each elevation

## 3.1 Analyses for each elevation

plot.list <- list()

for(site.id in substr(unique(COMMVARS$site),1,1)){

  print(site.id)

  SUB.VARS <- COMMVARS %>%
    filter(grepl(paste('^', site.id, sep = ""), site) & treatment %in% treats.to.consider)

  SUB.DATA <- COMMDATA[SUB.VARS$plotid, ]
  SUB.DATA <- decostand(SUB.DATA, method = 'hellinger')
  # Perform RDA and ANOVA and print the results
  rdamod <- rda(as.data.frame(SUB.DATA)~f+h+p+i+Condition(garden), 
                data = SUB.VARS)

  h <- how(blocks = SUB.VARS$garden)

  print(anova(rdamod, permutations = h, by = "terms"))

  # Goodness
  gdnsof <- goodness(rdamod, model = "CCA", proportional = T)
  gof2ax <- gdnsof[,1] + gdnsof[,2]
  cgof <- gof2ax[gof2ax > 0.2]
  cgof <- sort(cgof, decreasing = T)
  sign.names <- names(cgof[!is.na(cgof) & !is.infinite(cgof)])

  colors <- as.numeric(as.factor(SUB.VARS$treatment))

  # Alternative way to obtain these
  rda.dat <- list(sites = rdamod$CCA$wa[, c(1,2)],
                  species = rdamod$CCA$v[, c(1,2)],
                  biplot = rdamod$CCA$biplot[, c(1,2)])

  sites.dat <- as.data.frame(rda.dat$sites) %>%
    mutate(codes = rownames(rda.dat$sites),
           treats = ifelse(grepl("c", codes),"c",
                           ifelse(grepl("f", codes),"f", 
                                  ifelse(grepl("h", codes), "h", 
                                         ifelse(grepl("i", codes), "i","p")))),
           shapes = as.numeric(as.factor(treats)))

  species.dat <- as.data.frame(rda.dat$species) %>%
    mutate(names = rownames(as.data.frame(rda.dat$species)))

  arrow.dat <- as.data.frame(rda.dat$biplot) %>%
    mutate(treats = rownames(as.data.frame(rda.dat$biplot)))

  if(site.id == "y"){
    arr.col <- c("red","black","black","red")
    arr.lwd <- c(0.5,0.1,0.1,0.5)
    arr.lty <- c(1,2,2,1)
  } else {arr.col <- c("black","black","black","red")
  arr.lwd <- c(0.1,0.1,0.1,0.5)
  arr.lty <- c(2,2,2,1)
  }

  p1 <- ggplot(sites.dat,aes(x = RDA1, y = RDA2))+
    geom_point(aes(colour = treats), pch = sites.dat$shapes+14) + 
    geom_hline(yintercept = 0, lty = 2, alpha = 0.5) + 
    geom_vline(xintercept = 0, lty = 2, alpha = 0.5) + 
    geom_segment(data = arrow.dat, aes(x = 0, y = 0, 
                                       xend = RDA1, yend = RDA2),
                 arrow = arrow(length=unit(0.30,"cm"), 
                               ends="last", 
                               type = "closed"),
                 alpha = 0.8, 
                 col = arr.col,
                 lwd = arr.lwd,
                 lty = arr.lty)+
    geom_text(data = arrow.dat, aes(x = RDA1*1.1, y = RDA2*1.1, 
                                    label = treats),
              cex = 5)+
    theme_minimal()

  # scalar = 2.5

  p2 <- ggplot(species.dat, aes(x = RDA1*1.1, y = RDA2*1.1)) + 
    geom_text(data = species.dat %>%
                filter(names %in% sign.names), aes(label = names))+
    geom_segment(data = arrow.dat, aes(x = 0, y = 0, 
                                       xend = RDA1, yend = RDA2),
                 arrow = arrow(length=unit(0.30,"cm"), 
                               ends="last", 
                               type = "closed"),
                 alpha = 0.8, 
                 col = arr.col,
                 lwd = arr.lwd,
                 lty = arr.lty)+
    geom_text(data = arrow.dat, aes(x = RDA1*1.1, y = RDA2*1.1, 
                                    label = treats))+
    theme_minimal()

  plot.list[[site.id]] <- list(sites.ord = as_grob(p1),
                               species.ord = as_grob(p2))

}

## 3.2 Plot the ordination ----
ggpubr::ggarrange(plot.list[["w"]]$sites.ord,
                  plot.list[["w"]]$species.ord,
                  plot.list[["n"]]$sites.ord,
                  plot.list[["n"]]$species.ord,
                  plot.list[["y"]]$sites.ord,
                  plot.list[["y"]]$species.ord, ncol = 2, nrow = 3)

```

### Bray-Curtis dissimilarity

BC dissimilarity at the control plots was higher in Wanang than in Numba. Addition of herbivores created a monotonic decrease in dissimilarity with elevation (Fig. 4, reduction of dissimilarity means higher predictability of the plant community). In Wanang species composition changed mainly through turnover of species, in contrast to two other sites, where community changed through changes in the abundance structure (gradient component). 1. Baseline (?) species turnover is highest in Wanang. Gradient changes are smaller in wanang than in Numba. 2. Fungicide did not change the character of the community structure and assembly (patterns are similar in the case of both control and fungicide treatment). 3. Herbivore addition monotonically reduced dissimilarity (plots within the treatment were more similar to each other, reduced unpredictability of succession) by reducing it substantially in Yawan. This was caused by an increase in balanced component in numba and its reduction in Yawan. This means that species turnover between sites was reduced by additional herbivorous insects at the highest elevation. However, in both of these sites balanced component was similar. It was high in W and did not change there. 4. Insect removal monotonically reduces plant species turnover. Does not change the pattern of dissimilarity between elevations but substantially increases gradient component in Yawan (it increased slightly gradient component in Wanang and reduced significantly in Numba). Insect removal significantly reduced the species turnover (balanced component) in Yawan, and increased it in Numba. This means that the presence of insects (opposite of what the treatment does) increases species turnover in Y and reduces species turnover in N. Simultaneously it increase gradient component in Wanang and reduces it in Numba. Interestingly in Yawan further increase of herbivore pressure has the opposite effect again it brings it back to the previous level. Having additional herbivores has the same effect as not having insects at all in case of the balanced component. Predator treatmen is the same. In case of the gradient component only removal of insects was able to change it as indicated above. And only in wanang and numba. 5. Top-predators increase balanced and gradient component similar to insecticide treatement in Numba. Similarly to H and I it reduces balanced component in Yawan. But has no effect on BC dissimilarity. In Numba also reduce gradient component but does not do that in Yawan. Between sites: Balanced component is lowest in Yawan general BC dissimilarity is lowest in Yawan and gradient in Yawan is higher in Yawan than in Numba but similar to that in Wanang.

```{r}

# NEEDS TO BE FIXED

# library(betapart)
# 
# csites <- grepl("_c\\b", COMMVARS$plotid)
# fsites <- grepl("_f", COMMVARS$plotid)
# isites <- grepl("_i", COMMVARS$plotid)
# hsites <- grepl("_h", COMMVARS$plotid)
# psites <- grepl("_p", COMMVARS$plotid)
# 
# wsites <- COMMVARS$site == "wanang"
# nsites <- COMMVARS$site == "numba"
# ysites <- COMMVARS$site == "yawan"
# 
# # B-C dissimilarities in C plots between sites
# cdatc <- COMMDATA[grepl("c$", rownames(COMMDATA)), ]
# cvars <- COMMVARS[rownames(cdatc), ]
# 
# # Make a matrix
# for(i in 1:dim(cdatc)[1]){
#   
#   # Compare all other sites to the current i
#   for (j in (1:dim(cdatc)[1])[-i]) {
#     val <- beta.pair.abund(cdatc[c(i,j), ])
#   }
#   
# }
# 
# # Comparisons between
# wc.beta <- beta.pair.abund(COMMDATA[COMMVARS[wsites & csites,]$plotid,  ])
# 
# beta.df <- data.frame()
# 
# for(siteid in c("w","y","n")){
#   for(treat in c("_c\\b","_f","_h","_i","_p")){
#     
#     # Subset the data
#     nms <- rownames(COMMDATA)
#     sel.nms <- nms[grepl(siteid, nms) & grepl(treat, nms)]
#     site.treat.comm <- COMMDATA[sel.nms,]
#     
#     print(rownames(site.treat.comm))
#     
#     # Calculate beta diversity
#     subbeta <- beta.pair.abund(site.treat.comm)
#     
#     # Add to a data frame
#     balanced.df <- data.frame(site = siteid,
#                               treat = treat,
#                               vals = as.numeric(subbeta$beta.bray.bal),
#                               component = "balanced")
#     
#     gradient.df <- data.frame(site = siteid,
#                               treat = treat,
#                               vals = as.numeric(subbeta$beta.bray.gra),
#                               component = "gradient")
#     
#     bray.df <- data.frame(site = siteid,
#                               treat = treat,
#                               vals = as.numeric(subbeta$beta.bray),
#                               component = "bray")
#     
#     sub.df <- rbind(balanced.df, gradient.df, bray.df)
#       
#     # Merge the datset
#     beta.df <- rbind(beta.df,sub.df)
#     
#   }
# }
# beta.df
# 
# # Fix the site order
# beta.df$fsite <- factor(beta.df$site, levels = c("w","n","y"))
# 
# # Transform 0s to 0.001 and 1s to 0.999
# beta.df$trans.vals = ifelse(beta.df$vals == 1, 0.999999, 
#                             ifelse(beta.df$vals == 0, 0.000001, beta.df$vals))
# 
# # Tests for individual facets
# 
# library(lsmeans)
# library(emmeans)
# library(multcomp)
# library(multcompView)
# 
# cld.dat = data.frame()
# prws.comp <- list()
# 
# # CLD data for comparisons between elevations
# for (comp in unique(beta.df$component)){
#   print(comp)
#   for (trt in unique(beta.df$treat)){
#     print(trt)
#     
#     test.df <- beta.df %>%
#       filter(treat == trt,
#              component == comp)
#     
#     # print(test.df)
#     
#     mod1 <- betareg::betareg(trans.vals ~ fsite, data = test.df)
#     
#     lsmod1 <- lsmeans(mod1,
#                       pairwise ~ fsite,
#                       adjust="tukey")
#     
#     groups <- cld(lsmod1, Letters = c("abcdefg"))
#     
#     y_loc = test.df %>%
#       group_by(fsite) %>%
#       summarise(y_loc = mean(trans.vals))
#     
#     sub.cld <- data.frame(component = comp,
#                treat = trt,
#                fsite = groups$fsite,
#                gr.label = groups$.group,
#                y_loc = y_loc$y_loc[order(groups$fsite)])
#     
#     print(y_loc)
#     print(sub.cld)
#     
#     prws.comp[[paste(comp,trt)]] <- lsmod1
#     
#     cld.dat = rbind(cld.dat, sub.cld)
#     
#   }
# }
# 
# # cld.dat = data.frame()
# # prws.comp <- list()
# 
# # For data for comparisons between treatments we need pairwise comps 
# # (control vs treat) - this can be obtained from summary of the betareg
# 
# # library(glmmTMB)
# 
# pairwise.comp.res <- data.frame()
# for (comp in unique(beta.df$component)){
#   
#   for (ste in unique(beta.df$fsite)){
#     print(comp)
#     print(ste)
#     
#     test.df <- beta.df %>%
#       filter(fsite == ste,
#              component == comp)
#     
#     # print(test.df)
#     
#     mod1 <- betareg::betareg(trans.vals ~ treat, data = test.df)
#     
#     breg.table <- summary(mod1)$coefficients$mean
#     breg.pvals <- breg.table[,4]
#     
#     y_loc = test.df %>%
#       group_by(treat) %>%
#       summarise(y_loc = mean(trans.vals))
#     
#     pcr.row <- data.frame(comp = comp,
#                           fsite = ste,
#                           component = comp,
#                           treat = names(breg.pvals),
#                           p.vals = breg.pvals,
#                           y_loc = y_loc$y_loc,
#                           y_loc_trt = y_loc$treat)
#     
#     pcr.row <- pcr.row %>%
#       mutate(label = ifelse(p.vals <= 0.05 & treat != "(Intercept)", "*",""))
#     
#     pairwise.comp.res <- rbind(pairwise.comp.res,pcr.row)
#     
#   }
# }
# 
# pairwise.comp.res
# 
# # Plot
# # tiff(filename='figures/beta_part_panel.tif',
# #      height=5600,
# #      width=5600,
# #      units='px',
# #      res=800,compression='lzw')
# beta.part.plot <- ggplot(beta.df, aes(x = fsite, y = vals))+
#   geom_jitter(width = 0.1, alpha = 0.2, shape = 16, size = 1.8)+
#   stat_summary(fun.y = mean, na.rm = T,
#                geom = "point", size = 3, col = "gray30")+
#   stat_summary(fun.data = "mean_cl_boot", col = "gray30")+
#   stat_summary(fun.y = mean, aes(group = treat),
#                geom = "line", alpha = 0.8, col = "gray30", lty = 2)+
#   geom_text(data = cld.dat,
#             aes(x = fsite, y = y_loc + 0.5,
#                 label = gr.label),
#             size = 3)+
#   facet_grid(cols = vars(component), rows = vars(treat))
# beta.part.plot + theme_bw()
# 
# # Plot for the component comparisons
# pairwise.comp.res$fsite <- factor(pairwise.comp.res$fsite, levels = c("w","n","y"))
# 
# # tiff(filename='figures/beta_part_panel_treats.tif',
# #      height=5600,
# #      width=5600,
# #      units='px',
# #      res=800,compression='lzw')
# 
# # Alternative boxplot
# beta.part.plot <- ggplot(beta.df, aes(x = treat, y = vals))+
#   geom_boxplot()+
#   # stat_summary(fun.y = mean, na.rm = T,
#   #              geom = "point", size = 3, col = "gray30")+
#   # stat_summary(fun.data = "mean_cl_boot", col = "gray30")+
#   geom_text(data = pairwise.comp.res,
#             aes(x = y_loc_trt, y = y_loc + 0.5,
#                 label = label), cex = 6,  col = "red", lwd = 2)+
#   facet_grid(component ~ fsite)
# 
# beta.part.plot + theme_bw()

```

## Are changes random or deterministic for different treatments?

At the highest studied elevation (Yawan) all treatment caused a shift from composition being deterministic in their divergence (predictably unpredictable to random). Similar shift, but weaker in cases of insecticide and fungicide was observed for Wanang. Strikingly different assembly was observed for Numba, where communities showed rather random assembly and only addition of herbivores increased divergence of the community composition.

```{r}

all.pairs_max_accuracy <- read.csv("data/raup_crick/all.pairs_max_accuracy.csv")
all.pairs_max_accuracy$fsite <- factor(all.pairs_max_accuracy$fsite, 
                                       levels = c("w","n","y"), labels = c("Wanang",
                                                                           "Numba",
                                                                           "Yawan"))

test.df <- data.frame()
for (st in unique(all.pairs_max_accuracy$fsite)){
  for (cp in unique(all.pairs_max_accuracy$comp)){

    print(paste(st, cp))

    subdat <- all.pairs_max_accuracy %>%
      filter(fsite == st & comp == cp)

    mod <- nlme::lme(RC ~ treatment,
                     random = ~1|comp.id,
                     data = subdat)

    # print(t.test(RC ~ treatment, data = subdat, paired = TRUE))

    test.df <- rbind(test.df, data.frame(fsite = st,
                                         comp = cp,
                                         group1 = tolower(substr(cp,1,1)),
                                         group2 = tolower(substr(cp,6,6)),
                                         test.res = summary(mod)$tTable[2,5]))

    print(summary(mod))

  }
}

test.df <- test.df %>%
  mutate(label = ifelse(test.res < 0.05, "*"," "))

ggplot(all.pairs_max_accuracy, aes(x = treatment, y = RC))+
  # geom_boxplot()+
  ylim(c(-1.25,1.5))+
  stat_summary(fun.y = mean, na.rm = T,
               geom = "point", size = 4)+
  stat_summary(fun.data = "mean_cl_boot")+
  theme_bw()+
  ggtitle(paste("Accuracy = 0.001"))+
  geom_hline(yintercept = 0, lty = 2)+
  geom_hline(yintercept = 1, lty = 3)+
  geom_hline(yintercept = -1, lty = 3)+
  facet_grid(vars(fsite), vars(comp), scales = "free")+
  ggpubr::stat_pvalue_manual(data = test.df[test.df$label == "*",], label = "label",
                     y.position = 1.25, 
                     label.size = 6)

```

Making sense out of the above:

* Insects shape community composition at all elevations. Fungi affected community composition only in Yawan. Fungi increase the species richness, without affecting any other characteristics of the community (beta-diversity components). Whereas in Wanang despite having effect on richness and density the effects on the community composition was not visible possibly because of high baseline dissimilarity (variability in community composition) between sites. NUmba seems to be specific. Despite the effect being present they are being balanced in a way that left the community characteristics unchanged.

* Insect herbivores (P, I, H combined) can suppress plants in communities with low variability in species composition, which allows for new species to establish. Moreover, **insects reduce dissimilarity for highly variable communities** (is this true?). Exclusion of top predators has the same effect as insect exclusion on species turnover in Numba and Yawan. This is a probable result for a strong top-down control of the tri trophic interactions. Addition of generalist herbivores also correlated with the predator effects. If the effects of insects are mainly caused by specialists, then generalists interfering with their feeding could have the same results. But the effects can be more subtle than those of herbivore removal.

### Trait based analyses

In this section we focus on plant species traits and how these can help us predict the performance of individual species at a given plot. First we attempt to predict simply whether a species, considering it's traits, will be lost, gained or will remain in various treatment plots. We used random forest to investigate the importance of traits in predicting the species faith. Prediction accuracy and classifier_RF will tell us how good well the set of traits is (and which traits are most important in predicting faith) for the faith of individuals.

First we analyze only traits ignoring location. Later we can mak an attempt to see wether importance of

```{r}

source("code/pgls_data_prep.R")

subsetVars <- c("c", "i")

# 1. Prepare data ----
# Add richness
richdat <- data %>%
  group_by(site, garden, treatment) %>%
  summarise(richness = n()) %>%
  mutate(sgt = paste(site,garden,treatment, sep = "_"))

data.woody.brut <- data.woody.brut %>%
  mutate(sgt = paste(site,garden,treatment, sep = "_"))

dwbRich <- merge(data.woody.brut, richdat, by = "sgt")

# Filter and add estimated leaf area
dwbRichF <- dwbRich %>%
  mutate(est.leaf.area = num_leaf_weight * sla.w.m2kg1,
         lsla.d = log(sla.d.m2kg1),
         ltb = log(as.numeric(tot_bio)),
         labu = log(as.numeric(no_stems+1))) %>%
  dplyr::select(site.x, garden.x, treatment.x, 
         unified_names, tot_bio, no_stems, 
         perc.herb, water.cont, est.leaf.area, sla.d.m2kg1,
         richness, ltb, lsla.d, labu) %>%
  filter(water.cont > 0)

names(dwbRichF) <- c("site", "garden", "treatment", "unified_names","tot_bio",
                     "no_stems", "perc.herb", "water.cont", "est.leaf.area",
                     "sla.d.m2kg1", "richness", "ltb","lsla.d","labu")

# Subset a data-set to only control and treatment plots
subDat <- dwbRichF %>%
  filter(treatment %in% subsetVars)

lgsData <- data.frame()

for (st in unique(subDat$site)){
  for (gard in unique(subDat$garden)){

    stGardDat <- subDat %>%
            filter(site == st,
                   garden == gard)

    # Some quality control
    if(dim(stGardDat)[1] == 0){
      next
      }
    if(length(unique(stGardDat$treatment)) < 2){
      next
      }

    specInCtr <- unique(stGardDat[stGardDat$treatment == subsetVars[1], ]$unified_names)
    specInTrt <- unique(stGardDat[stGardDat$treatment == subsetVars[2], ]$unified_names)

    spLost <- specInCtr[!(specInCtr %in% specInTrt)]
    spGained <- specInTrt[!(specInTrt %in% specInCtr)]

    stGardDat$pg.cat <- ifelse(stGardDat$unified_names %in% spLost, "lost",
                               ifelse(stGardDat$unified_names %in% spGained, "gained", "stayed"))

    lgsData <- rbind(lgsData, stGardDat)

  }
}
```

I perform the sane analyses for each treatment.

```{r}

# Function that performs RF analysis for a given treatment

# It queries the lgsData that needs to be preapred using separate code (as above)

getLgsData <- function (comp.trt = "i", data) {

  # This function queries the data to compute log

}

# Returns:
# 1) Classifier RF table
# 2) Confusion matrix
# 3) Regression tree example graph

runRF <- function (lgsData) {

}

```

```{r}
# Random Forest
# Loading package
library(caTools)
library(randomForest)

rfData <- lgsData %>%
  mutate(pg.cat = as.factor(pg.cat),
         lest.leaf.area = log(est.leaf.area),
         lstem = log(no_stems)) %>%
  dplyr::select(pg.cat, lstem, perc.herb, water.cont,
                lest.leaf.area, richness, ltb, lsla.d, treatment)

# Try standardized, non-correlated variables
rfDataRF <- rfData %>%
  filter(!is.infinite(lest.leaf.area)) %>%
  dplyr::select(pg.cat, perc.herb, water.cont,
                richness, ltb, lsla.d)

# Standardized
rfDataRF[,-1] <- scale(rfDataRF[,-1])

# Splitting data in train and test data
split <- sample.split(rfDataRF, SplitRatio = 0.7)
train <- subset(rfDataRF, split == "TRUE")
test <- subset(rfDataRF, split == "FALSE")

classifier_RF = randomForest(x = train[-1],
                             y = train$pg.cat,
                             ntree = 500)

classifier_RF

# Predicting the Test set results
y_pred = predict(classifier_RF, newdata = test[-1])

# Variable importance plot
varImpPlot(classifier_RF)
```

Below is a confusion matrix for the RF

```{r, results='asis'}
# Confusion Matrix
confusion_mtx = table(test[, 1], y_pred)
print(knitr::kable(confusion_mtx))
```

Using collected variables we cannot predict the faith of an individual species with high certainty. We can only predict which plants can be observed (stayed) at both control and insecticide treated plots. Error of that prediction is relatively low (0.16). For this prediction total biomass is the most important. SLA is least important, together with the number of stems. To visualize the division we run a single regression tree.

```{r, fig.width = 25, fig.height= 10}

library(rpart)
library(rpart.plot)

# Regression C50
library(C50)
library(printr)

C50Fit <- C5.0(pg.cat ~., data=train,
               control = C5.0Control(minCases = 5))

C50pred <- predict(object=C50Fit, 
                   newdata=test, 
                   type="class")

plot(C50Fit, cex = 0.5)

```

Lets see how good of a job the regression tree did.

```{r, results='asis'}

print(knitr::kable(table(C50pred, test$pg.cat)))

```

The above tree is just a one example of hundreds of averaged trees used in the random forest. A few key conclusions can be drawn from the above analyses is that we cannot in certain predict the faith of a tree species. But its biomass is the best predictor together with water content. SLA seems to be the least important.

Let us now move to test SLA ability to predict the biomass of an individual. This is a bit tricky because we will use SLA estimated at the end of the experiment. From the leaf frame analyses that spanned all plant species in all treatment plots we saw high plasticity of that trait.

## Trait-based predictions of a species success

### At the level of the whole community

```{r}
# CWM vs richness
```

Community-weighted mean SLA (CWM SLA) was higher at low diversity plots. This means that species with RELATIVELY higher SLA were the ones that dominated the communities. In Numba baseline SLA was reduced even further. In Numba also fungicide treatment had lower baseline SLA.

In both cases slope of the relationship was increased (less steep, it was flatter, no relationship).

```{r, fig.cap='Community weighted mean vs number of species at a site for the studied treatments at three elevations. Size of a point indicates total plant biomass of that plot.'}

# SLA values of individual plant species for each treatment.

# I am using the data.woody.brut data-set from the pgls_data_prep script. It has
# filtered woody palnt names, cleaned sla.d values and limited sla range.

# Define CWM function
cwmFun <- function(sla.d.m2kg1, tot_bio){
  # summarize function: CWM for sites
  return(mean(sla.d.m2kg1*(tot_bio/sum(tot_bio, na.rm = T)), na.rm = T))
}

# siteSLA <- data.woody.brut %>%
#   filter(unified_names %in% trait.dat[trait.dat$Life.form == "woody", ]$dsName) %>%
#   mutate(tot_bio = ifelse(is.infinite(tot_bio), NA, tot_bio),
#          sla.d.m2kg1 = ifelse(is.infinite(sla.d.m2kg1), NA, sla.d.m2kg1)) %>%
#   group_by(site, garden, treatment) %>%
#   summarise(cwm.sla = cwmFun(sla.d.m2kg1, n.tot_bio),
#             sp.rich = n())

siteSLA <- data.woody.brut %>%
  filter(treatment %in% c('i','c','p','h', 'f')) %>%
  mutate(tot_bio = ifelse(is.infinite(tot_bio), NA, tot_bio),
         sla.d.m2kg1 = ifelse(is.infinite(sla.d.m2kg1), NA, sla.d.m2kg1)) %>%
  group_by(site, garden, treatment) %>%
  summarise(cwm.sla = cwmFun(sla.d.m2kg1, n.tot_bio),
            sp.rich = n(),
            total.bio = sum(n.tot_bio))

siteSLAdf <- as.data.frame(siteSLA)

rownames(siteSLAdf) <- paste(substr(siteSLA$site,1,1), 
                             siteSLA$garden, 
                             "_", 
                             siteSLA$treatment, sep = "")

# Plot of CWM vs species richness
require(scales)
p1 <- ggplot(siteSLAdf, aes(x = sp.rich, y = cwm.sla, color = treatment))+
  geom_point(aes(size = total.bio)) + 
  scale_x_continuous(trans = 'log2',
    breaks = trans_breaks("log2", function(x) 2^x),
    labels = trans_format("log2", math_format(2^.x))) +
  scale_y_continuous(trans = 'log2',
    breaks = trans_breaks("log2", function(x) 2^x),
    labels = trans_format("log2", math_format(2^.x))) +  theme_minimal()

p2 <- ggplot(siteSLAdf, aes(x = treatment, y = cwm.sla, fill = site))+
  geom_boxplot()+
  scale_y_continuous(trans = 'log2',
    breaks = trans_breaks("log2", function(x) 2^x),
    labels = trans_format("log2", math_format(2^.x))) +  theme_minimal()

p3 <- ggplot(siteSLAdf, aes(x = site, y = cwm.sla, fill = treatment))+
  geom_boxplot()+
  scale_y_continuous(trans = 'log2',
    breaks = trans_breaks("log2", function(x) 2^x),
    labels = trans_format("log2", math_format(2^.x))) +  theme_minimal()

ggpubr::ggarrange(p1,p2, p3)

# What is that value for the Y, g8, p?
# This seem to be an outlier - very low total biomass
# data.woody.brut %>% filter(site == 'yawan',
#                            garden == 'g8',
#                            treatment == 'p')

```
```{r}
# Test for the relationship and the differences
require(sjPlot)
cwm.lm <- lm(log(cwm.sla)~log(sp.rich)*treatment + site, data = siteSLAdf)
cwm.lm.nosite <- lm(log(cwm.sla)~log(sp.rich)*treatment, data = siteSLAdf)
cwm.lm.nosite.noint <- lm(log(cwm.sla)~log(sp.rich)+treatment, data = siteSLAdf)
cwm.lm.onlysp <- lm(log(cwm.sla)~log(sp.rich), data = siteSLAdf)

# cwm.glm <- glm(cwm.sla~sp.rich*treatment + site, data = siteSLAdf,
#              family = gaussian(link = 'log'))
# cwm.glm.nosite <- glm(cwm.sla~sp.rich*treatment, data = siteSLAdf,
#                     family = gaussian(link = 'log'))
# cwm.glm.nosite.noint <- glm(cwm.sla~sp.rich+treatment, data = siteSLAdf, 
#                           family = gaussian(link = 'log'))
# cwm.glm.onlysp <- glm(cwm.sla~sp.rich, data = siteSLAdf, family = gaussian(link = 'log'))

# tab_model(cwm.lm.nosite.noint)

# anova(cwm.lm.nosite, cwm.lm.onlysp)
knitr::kable(AIC(cwm.lm, cwm.lm.nosite, cwm.lm.nosite.noint, cwm.lm.onlysp))
# knitr::kable(AIC(cwm.glm, cwm.glm.nosite, cwm.glm.nosite.noint, cwm.glm.onlysp,
                 # cwm.lm, cwm.lm.nosite, cwm.lm.nosite.noint, cwm.lm.onlysp))

# tab_model(cwm.glm)
tab_model(cwm.lm.onlysp)

# plot(cwm.glm)
# plot(cwm.lm)

# there are no differences

```

The CWM and species number relation is strong and does not change between the treatments and sites.

### Individual species performance

Our original idea was that SLA (approximation of a growth rate) should become a better predictor of plant success (fitness, measured as total biomass in a plot, or positive change in biomass - deltaBio) with elevation.

## SLA

### Can SLA predict change in composition?

SLA doesn’t predict RDA position along C->I, C->H, C->P, C -> F axes, at any elevation.

```{r}

# Figure

```

Community composition change under a treatment may be independent of SLA. [key indicator of functionally important aspects of photosynthetic capacity, and is positively correlated with ecosystem productivity (White et al., 2000, Madani et al., 2017, Anderson et al., 2020). Reflects the abilities of light capture and adaption to environments of plants (Gao, Wang, and Zhang 2022)]

### Can SLA predict 'success' of a plant species?

In control plots with increasing elevation we hypothesize that SLA (used here as a proxy for the growth rate) should predict species' success (fitness, dominance, change in biomass in response to the treatment) better at higher elevation. Specifically, we expect that plants with higher SLA should be characterized by greater ability to dominate the community (higher total biomass - or should it be relative biomass? Or should it be biomass change?) in treatments where natural enemies were removed. We assume that in these cases inter-specific plant competition should be driven mainly by growth rates.

However there are a few contingent effects that make interpretation of this relationship problematic:

1. SLA of a species may be *determined by the community in which it resides*. CWM and richness dependency points to that. However, at the *individual level we do not see that species generally change their SLA in response to richness of the plot* on which they reside. This suggests that increased CWM SLA at low richness plots is driven by dominance of high SLA plants rather than changes in SLA by the same dominants, as would be suggested by the lack of differences in the community composition. This lack of significant effect of treatments in the RDA analysis may in turn be caused by a large variation between gardens (high variability in the control sites).

```{r}

# SLA of selected species at different treatments.

# Sla of selected species vs species richness

```

2. From our data we see that at some treatments same species may have different average SLA

3. Change in biomass (delta biomass) may be a better measurement of a species' response to a treatment. This excludes also species that were lost or gained as a response to a treatment. However, it was shown in the previous analysis that these plants were rather small and either in poor condition or just got established at a plot.

4. Change in biomass may be correlated to change in SLA (LRR like plots: what are the results of this?)

```{r}

# Change in SLA vs change in Biomass

```

5. Biomass/$\Delta$Biomass may be smaller if there are many individuals of the same species around. Therefore I think we should **include also the effect of the co-specifics’ abundance**.

6. Additionally total biomass may influence the biomass change, i.e, change in biomass may be smaller for smaller plants and larger for larger plants. What does this measure mean? Delta BIO measures how dominance has changed. If there was a dominant plant in the control and its biomass decreased a lot in the treatment what that would mean?

Taking the above considerations into account we formulated the final statistical model. The response variable was the $\Delta$Biomass. We tested whether it is dependent on SLA but also include the corresponding *site richness* and the *abundance of co-specifics* as co-variate. To test treatment influence on $\Delta$Biomass vs SLA relationship we included the interaction between SLA and treatment variable. We repeated model for each site and evaluated the model performance to compare power of SLA to predict the biomass changes at each elevation. We added a quadratic term for the Phylogenetic correlations are preserved:

#### Complex model model
$\Delta$Biomass ~ 
SLA $\times$ site + SLA^2 $\times$ site + 
water.content $\times$ site + water.content^2 $\times$ site + 
perc.herb $\times$ site  + perc.herb^2 $\times$ site + 
richness + abundance + $\Delta$SLA + 
(Phylogenetic structure).

Random effect is not necessary because we have paired comparisons of control vs treatment within the same garden. **Another option** is to include site as a random factor and get rid of all the interaction terms.

This model needs to be applied separately for each treatment comparison.

First we analyse the focused model, where we use only SLA as a predictor of species' success. Further we compare the models' performance between sites to explore the idea that SLA is an increasingly better predictor of a species responses with elevations.

#### 1.1 Insecticide

```{r}
# Create data-set for a given treatment
deltaData <- data.frame()

# Change total bio variable into numeric
lgsData$tot_bio <- as.numeric(lgsData$tot_bio)

# Loop through each species in each garden and site and calculate their deltaBio
for (st in unique(lgsData$site)){

  for (gard in unique(lgsData$garden)){

    stGardDat <- lgsData %>%
      filter(site == st,
             garden == gard)

    if (dim(stGardDat)[1] == 0){
      # If subset is empty skip to another one
      next
    }

    if (length(unique(stGardDat$treatment)) < 2){
      # If there is species in only one treatment skip to another one
      next
    }

    specInCtr <- unique(stGardDat[stGardDat$treatment == subsetVars[1], ]$unified_names)
    specInTrt <- unique(stGardDat[stGardDat$treatment == subsetVars[2], ]$unified_names)
    specInBoth <- specInCtr[specInCtr %in% specInTrt]

    cDat <- stGardDat %>%
      filter(unified_names %in% specInBoth,
             treatment == subsetVars[1]) %>%
      arrange(unified_names)
    tDat <- stGardDat %>%
      filter(unified_names %in% specInBoth,
             treatment == subsetVars[2]) %>%
      arrange(unified_names)

    # Calculate LRR for biomass and SLA change

    cDat$comparison <- subsetVars[2]

    if (subsetVars[2] == "h"){
      cDat$lrrtb <- log(tDat$tot_bio/cDat$tot_bio)
      cDat$lrrsla <- log(tDat$sla.d.m2kg1/cDat$sla.d.m2kg1)
    } else{
      cDat$lrrtb <- log(cDat$tot_bio/tDat$tot_bio)
      cDat$lrrsla <- log(cDat$sla.d.m2kg1/tDat$sla.d.m2kg1)
    }

    deltaData <- rbind(deltaData, cDat)

  }
}

# Log transform estimated leaf area
deltaData <- deltaData %>%
          mutate(lest.leaf.area = log(est.leaf.area))%>%
  filter(!is.infinite(lest.leaf.area))

# The Focused model
# Only the predictive power of SLA, water content and percentage herbivory on deltaBio

# Wanang ----
wdata <- deltaData %>% 
  filter(unified_names %in% myTree.rand.poly$tip.label) %>%
  filter(site == "wanang")

dd.gls.focsed.w <- nlme::gls(lrrtb ~ lsla.d,
                     data = wdata,
                     method = "ML",
                     correlation = corMartins(2.718282, 
                                              phy = myTree.rand.poly, 
                                              form = ~unified_names,
                                              fixed = T),
                     na.action = na.omit)

# Model results

# Predict values
pred.vals <- AICcmodavg::predictSE.gls(dd.gls.focsed.w, wdata, 
                                       se.fit = T)
wdata <- wdata %>%
  mutate(preds = pred.vals$fit,
         se.preds = pred.vals$se.fit,
         se.upr = preds + se.preds,
         se.lwr = preds - se.preds)

# print the MSE
mse_focused_df <- data.frame(mse = with(wdata, mean((lrrtb - preds))^2),
                             treatment = "Insecticide", site = "Wanang")

# Numba ----
ndata <- deltaData %>% 
  filter(unified_names %in% myTree.rand.poly$tip.label) %>%
  filter(site == "numba")

dd.gls.focsed.n <- nlme::gls(lrrtb ~ lsla.d,
                     data = ndata,
                     method = "ML",
                     correlation = corMartins(2.718282, 
                                              phy = myTree.rand.poly, 
                                              form = ~unified_names,
                                              fixed = T),
                     na.action = na.omit)

# Predict values
pred.vals <- AICcmodavg::predictSE.gls(dd.gls.focsed.n, ndata, 
                                       se.fit = T)
ndata <- ndata %>%
  mutate(preds = pred.vals$fit,
         se.preds = pred.vals$se.fit,
         se.upr = preds + se.preds,
         se.lwr = preds - se.preds)

# print the MSE
mse_focused_df <- rbind(mse_focused_df,  
                        data.frame(mse = with(ndata, mean((lrrtb - preds))^2),
                                   treatment = "Insecticide", site = "Numba"))

# Yawan ----
ydata <- deltaData %>% 
  filter(unified_names %in% myTree.rand.poly$tip.label) %>%
  filter(site == "yawan")

# The model
dd.gls.focsed.y <- nlme::gls(lrrtb ~ lsla.d,
                     data = ydata,
                     method = "ML",
                     correlation = corMartins(2.718282, 
                                              phy = myTree.rand.poly, 
                                              form = ~unified_names,
                                              fixed = T),
                     na.action = na.omit)

# Predict values
pred.vals <- AICcmodavg::predictSE.gls(dd.gls.focsed.y, ydata, 
                                       se.fit = T)
ydata <- ydata %>%
  mutate(preds = pred.vals$fit,
         se.preds = pred.vals$se.fit,
         se.upr = preds + se.preds,
         se.lwr = preds - se.preds)

# print the MSE
mse_focused_df <- rbind(mse_focused_df,  
                        data.frame(mse = with(ydata, mean((lrrtb - preds))^2),
                                   treatment = "Insecticide", site = "Yawan"))
```

```{r, results='asis'}
library(sjPlot)

tab_model(dd.gls.focsed.w, dd.gls.focsed.n, dd.gls.focsed.y)

```

```{r, fig.cap = "Empirical and predicted values of a simple model of biomass change vs logarithm of SLA for each studied sites."}

# Plot the relationship and add predicted values
wpredPlot <- ggplot(wdata, aes(y = lrrtb, x = lsla.d))+
  geom_point() +
  geom_line(data = wdata, aes(x = lsla.d, y = preds), lwd = 1,
            lty = 2)+
  # geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15) +
  ggtitle("Wanang 200 m a.s.l.")

npredPlot <- ggplot(ndata, aes(y = lrrtb, x = lsla.d))+
  geom_point() +
  geom_line(data = ndata, aes(x = lsla.d, y = preds), 
            lwd = 1, lty = 2)+
  # geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15) + 
  ggtitle("Numba 750 m a.s.l.")

ypredPlot <- ggplot(ydata, aes(y = lrrtb, x = lsla.d))+
  geom_point() +
  geom_line(data = ydata, aes(x = lsla.d, y = preds), lwd = 1)+
  geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15) + 
  ggtitle("Yawan 1900 m a.s.l.")

ggpubr::ggarrange(wpredPlot, npredPlot, ypredPlot,
                  nrow = 1)

```

```{r}
# Plot the relationship and add predicted values
ggplot(rbind(ydata, ndata, wdata), aes(y = lrrtb, x = lsla.d, color = site))+
  geom_point() + 
  geom_line(data = rbind(ydata, ndata, wdata), 
            aes(x = lsla.d, y = preds), lwd = 1)+
  geom_ribbon(aes(ymin=se.lwr, ymax=se.upr), alpha = .15)

sjPlot::tab_df(mse_focused_df)

```

The above shows that focused model, where we used SLA as the only predictor of the plant change in biomass has a declining ability to predict change in biomass at the insecticide plot.