### 2023-07-19 ms1_supplement.Rmd
Models for LRR

---
title: "Supplement X. Bottom-up drivers of the LRR effects"
author: "Piotr Szefer"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cars}
source('code/lrr_and_community_chars_correlations.R')
```

```{r}
tab_model(biolrr.mod0)
```

```{r pressure, echo=FALSE, fig.width=10, fig.height=10}
ggpubr::ggarrange(p1, p1.1,
                  nrow = 2)
```